---
title: "Time Series PR to EIR" 
subtitle: "Computing a fitted EIR time series from PR time series data" 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: 
    theme: paper 
vignette: >
  %\VignetteIndexEntry{Time Series PR to EIR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r}
library(knitr)
library(ramp.xds)
library(ramp.work)
```

```{r, echo=F}
#devtools::load_all()
```

# 

A dynamical systems model, $\mathcal M$, describing malaria... 

**Step-Ahead Fitting** 

Suppose we have a time series with observations:

$$x_1, x_2, x_3, \ldots, x_n$$ 

at time points

$$t_1, t_2, t_3, \ldots, t_n$$ 

we are looking for a set of ar values: 

$$\alpha_1, \alpha_2, \alpha_3, \ldots, \alpha_n$$

Let $$F_{\mathcal M}(t, \alpha_t, y_t, {\mathcal M}) = y_{t+1}$$ be a function that solves (updates) the model, $\mathcal M$ once at time $t$ and returns the state variables $y$ 

The pseudo-code for the fitting algorithm is: 

+ initialize the model, $\mathcal M$, such that 

	+ $F_{\mathcal M}(t_1, \alpha_1, y_1, {\mathcal M}) = y_1$ 
	
	+ $F_{obs}(y_1) = x_1$
	
+ Iteratively compute $\alpha_2$ such that: 

	+ $F_{\mathcal M}(t, \alpha_t, y_t, {\mathcal M}) = y_{t+1}$ 
	
	+ $F_{obs}(y_{t+1}) = x_{t+1}$

## Verification

This function probably needs to get pushed into the code somehow. 

```{r, show_pr}
show_pr = function(model, get_pr=F_pr, i=1){
  y = get_inits(model, flatten=TRUE)
  get_pr(list_Xvars(y, model, 1), model$Xpar[[1]])
}
```

We set up a model:

```{r}
SI <- xds_setup(Xname = "SIS", MYZname = "SI")
```

Note that the initial conditions are not what we wanted: 

```{r, eval=F}
show_pr(SI)
```

We initialize the model to have a *Pr*PR 

```{r, eval=F}
SI <- dts_init_by_pr(.4, SI, F_pr, 365)
show_pr(SI)
```

We find the attack rate at the steady state:   

```{r, eval=F}
ar <- dts_pr2ar_step(.4, si, F_pr)
ar
```

and we generate a time series:

```{r, eval=F}
set.seed(230)
ar_ts <- ar*c(1, rlnorm(14, 1, 1))
ar_ts
```

We use it to create a *Pf*PR time series:

```{r, eval=F}
si1 = si
pr_ts = show_pr(si1)
for(i in 2:length(ar_ts)){
  y = update_by_ar(ar_ts[i], si1, 1)
  si1 = update_inits(y, si1)
  pr_ts = c(pr_ts, show_pr(si1))
}
pr_ts
```

Now, we pass the *Pf*PR time series to `dts_pr2ar_ts` and test that it returns our AR time series:  

```{r, eval=F}
ar_t <- dts_pr2ar_ts(pr_ts, si, F_pr)
sum((ar_ts-ar_t)^2) < 1e-9
```

