---
title: "History" 
subtitle: "Modeling the History of Exposure" 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document 
vignette: >
  %\VignetteIndexEntry{History}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=F}
library(ramp.xds)
library(ramp.work)
```


```{r, echo=F}
#devtools::load_all()
```

In malaria, the immune status of the population depends on the history of exposure. Here, we develop a method to estimate the history of exposure from a time series describing the *Pf*PR in a population over time. The algorithm is called `pr2eir_history` that fits the signal. 

## Pseudocode

1. Make an initial guess about the average annual EIR. 

2. Set initial conditions for the model to the steady state using the average EIR from the first six months of data.  

3. Choose a function family, and fit the seasonal pattern: 

    a. Given an initial seasonality pattern, fit the phase.  
    
    b. Fit the amplitude.

4. Fit a spline function to describe inter-annual variability. 

5. At this point, the model should be reasonably close to the data, but we might want to adjust:

    a. Refit the average annual EIR
    
    b. Refit the phase 
    
    c. Refit the amplitude 
    
    c. Refit the trend 
    
6. Return a model, fitted to the data
    
## Example 

```{r}
set.seed(127)
F_s1 <- make_function(makepar_F_sin(bottom=1))
F_t1 <- make_function(makepar_F_spline(365*(0:5), c(.7, 2, 1.5, .9, .6, .8)))
tt <- seq(0, 365*5, by = 30)
pr <- .28*F_s1(tt)*F_t1(tt)
obs_pr <- rnorm(length(tt), pr, .1) 
plot(tt/365, obs_pr, type = "b", xlab = "Time (in years)", ylab = "PR", ylim = c(0,1))
lines(tt/365, pr, lwd=2) 
```


The first step is to set up the model we want to use. This uses the default **X** component model, `SIS,` and we set up the model using the built-in sin function family from `ramp.xds`: 

```{r shell}
mod  <- xds_setup_cohort(season_par = makepar_F_sin())
```

To call the function, we pass the observed pr time series, the times, and the model:  

```{r pr2eir_history}
mod1 <- pr2eir_history(obs_pr, tt, mod, twice=FALSE)
```

```{r}
Fx <- mod1$EIRpar$F_trend
integrate(Fx,min(tt), max(tt))$val/(max(tt)-min(tt)) 
```


```{r}
mod2 <- pr2eir_history(obs_pr, tt, mod, twice=TRUE)
```

We can visualize the results: 

```{r vis}
ttn = seq(-365, max(tt), length.out=600)
xds_plot_PR(mod1)
xds_plot_PR(mod2, clrs="blue", add=TRUE)
lines(tt, obs_pr, xlab = "Time (in years)", ylab = "PR")
lines(tt, pr, col = "red")  
```

