---
title: "Metric Conversion" 
subtitle: "" 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document
vignette: >
  %\VignetteIndexEntry{Metric Conversion}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

To calibrate models, a user might know the value of one term, but they need to set the value of another. **`ramp.work`** includes a set of functions that do metric conversion. These functions require `xds_scaling` 

## `xde_pr2eir()`

After running `xds_scaling,` the outputs `xds_obj$outputs$scaling` has stored values of the EIR and the PR. The function `xde_pr2eir()` uses linear interpolation. 


## Example 

```{r, message=FALSE}
library(ramp.xds)
library(ramp.work)
devtools::load_all()
```

The algorithms require the outputs of `xds_scaling`

```{r, fig.height=4, fig.width=7}
Spar  <- makepar_F_sin(bottom = 0.2, pw=2)
model <- xds_setup_eir(eir = 1/365, season_par = Spar)
model <- xds_scaling(model)
clrs <- viridisLite::turbo(length(model$scaling$eir))
xds_plot_eirpr(model, clrs=clrs)
add_eirpr_orbits(c(8, 12, 16,20), model, clrs=clrs)
```

We can create 50 randomly chosen values of the *Pf*PR, and output the associated *Pf*EIR values

```{r}
prs = c(0.001, runif(25, 0, 1), 0.999)
preir_i = xds_pr2eir(prs, model)
```

The function flags any values that are outside of the acceptable range. This may not seem important for the SIS model, but the range of other models can be bounded, so we don't want to return nonsense values.  

```{r}
preir_i$errors
```

We can plot the others:  

```{r}
xds_plot_eirpr(model)
with(model$scaling, points(aeir, pr, pch = 15))
with(preir_i, points(365*eir, pr, pch = 19, col = "red"))
```
