---
title: "Pipeline"
output: html_document
---

We want to automate the process of model fitting and baseline reconstruction. It is a common enough task to spell it out. 

```{r}
library(ramp.xds)
library(ramp.library)
library(ramp.control)
library(ramp.uganda)
```

```{r, echo=F}
devtools::load_all()
```

## Working Directory

```{r}
save_to = "/Users/smitdave/git/ramp.work/vignettes/"
#save_to = "/Users/smitdave/git/ramp.uganda/models/"
#save_to = "/Users/smitdave/Library/CloudStorage/Box-Box/RAMP/models/" 
```


## Base Model 

We want to work with a model family, so we set one up. We will want to use the `eirpr` table created by `xde_scaling_lambda`, but we only need to do it once: 

```{r, eval=F}
# The block sets `eval=F` to avoid having to repeat this step
x0 <- Sys.time()
sip_si <- xds_setup(MYZname = "SI", Xname = "SIP")
sip_si <- xde_scaling_lambda(sis_si)
Sys.time()-x0
```

```{r, eval=F}
save_as <- "sis_si.rds" 
file_name <- paste(save_to, save_as, sep="")
print(file_name)
```

```{r, eval=F}
file_name <- paste(save_to, save_as, sep="")
saveRDS(sip_si, file_name)
```

```{r, eval=F}
sis_si_base <- readRDS(file_name)
```

## Read Data 


```{r prts}
i = 132
prts <- get_district_pfpr_i(i)
district_name <- prts$district_name[1]
```


```{r viz}
with(prts, {
  plot(jdate, pfpr, ylim = c(0,0.65), main = "Terego District", xlab = "Julian Date (day 1 is Jan 1, 2015)", ylab = "PR") 
  lines(jdate, pfpr)
})
```
## Naive History

```{r}
devtools::load_all()
```

```{r, eval=F}
sis_si_base <- readRDS(file_name)
sis_si_base$Lpar[[1]]$Lambda
```

```{r, eval=F}
sis_si_base$Lpar[[1]]$season_par <- makepar_F_sin(bottom=1, phase=180) 
sis_si_base$Lpar[[1]]$F_season <- make_function(makepar_F_sin(bottom=1,phase=180))
```


```{r, eval=F}
fit_mod <- setup_fitting(sis_si_base, prts$pfpr, prts$jdate)

fit_mod$Lpar[[1]]$Lambda
```

```{r, eval=F}
fit_mod <- xds_solve(fit_mod, times = prts$jdate)
profile_pr(fit_mod, prts, y1=TRUE)
```

```{r, eval=F}
fit_mod1 <- fit_mean(prts$pfpr, prts$jdate, fit_mod)
fit_mod1$Lpar[[1]]$Lambda
```




```{r, eval=F}
fit_mod <- last_to_inits(fit_mod) 
profile_pr(fit_mod, prts, y1=TRUE)
```

```{r}
devtools::load_all()
```

```{r, eval=F}
t0 <- Sys.time()
fit_mod1 <- fit_mean(prts$pfpr, prts$jdate, fit_mod)
Sys.time()-t0
```


```{r, eval=F}
t0 <- Sys.time()
fit_mod1 <- pr2history(prts$pfpr, prts$jdate, fit_mod)
Sys.time()-t0
```
```{r, eval=F}
fit_mod1 <- xds_solve(fit_mod1, times = prts$jdate)
profile_pr(fit_mod1, prts, y1=TRUE)
```

```{r, eval=F}
t0 <- Sys.time()
fit_mod2 <- pr2history(prts$pfpr, prts$jdate, fit_mod, twice=TRUE)
Sys.time()-t0
```
## Setup Vector Control
```{r}
devtools::load_all()
```

```{r, eval=F}
irs_history <- make_irs_history(district_name)
bednet_history <- make_bednet_history(district_name)
```

```{r}
devtools::load_all()
```

```{r, eval=F}
recon <- setup_vector_control_evaluation(fit_mod, irs_history, bednet_history)
```

```{r, eval=F}
recon$fitting$data$modified
```

## Setup Imputation 

```{r, eval=F}
recon <- setup_imputation(recon)
```

## Setup Imputation 

## The Pipeline

This is a pipeline for Uganda. 

```{r}
#' @title Reconstruction Pipeline 
#' @description 
#' Run the baseline reconstruction pipeline for the 
#' \eqn{i^{th}} 
#' 
#' @param i the index of a district to pull from  
#' @param base_model_name the stem name of a model   
#' @param wd_name the working directory name 
#' @param impute 
#'
#' @returns
#' @export
#'
#' @examples
pipeline_i = function(i, base_model_name, wd_name, impute="mean"){
  # read the model from 
  model_file <- paste(wd_name, base_model_name, ".rds", sep="")
  model <- readRDS(model_file)
  print(model_file)
  
  # pull a time series  
  prts <- get_district_pfpr_i(i)
  district_name <- prts$district_name[1]
  dir_name <- prts$dir_name[1]
  
  # setup the fitting 
  fit_mod <- setup_fitting(model, prts$pfpr, prts$jdate)
  fit_mod <- pr2history(prts$pfpr, prts$jdate, fit_mod)
  filename1 <- paste(base_model_name, "_", dir_name, "_naive_history.rds", sep="") 
  saveRDS(fit_mod, paste(wd_name, filename1, sep=""))
  print(filename1)
  
  irs_history <- ramp.uganda::make_irs_history(district_name)
  bednet_history <- ramp.uganda::make_bednet_history(district_name)
  recon <- setup_vector_control(model, irs_history, bednet_history)
 
  recon <- setup_baseline_reconstruction(recon,)
  
  recon <- reconstruct_baseline(fit_mod, prts$pfpr, prts$jdate) 
  recon <- fit_mod
  filename2 <- paste(base_model_name, "_", dir_name, "_reconstructed_history.rds", sep="") 
  saveRDS(recon, paste(wd_name, filename2, sep=""))
  print(filename2)
  
  baseline <- fit_mod 
  filename3 <- paste(base_model_name, "_", dir_name, "_baseline.rds", sep="") 
  saveRDS(baseline, paste(wd_name, filename3, sep=""))
  print(filename3)
  

  return(list(district_name = district_name, 
              dir_name = dir_name, 
              prts = prts, 
              base_model = model,
              naive_history = filename1, 
              reconstructed_history = filename2,
              baseline = filename3)) 
}

```

this is ugly 

```{r, eval=F}
i = 132
base_mod = "sip_si_base" 
wd = "/Users/smitdave/git/ramp.uganda/models/"

pipeline_i(i, base_mod, wd, impute="mean") -> donkey_doug
```
