---
title: "Scaling for Malaria Metrics" 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: 
    theme: paper 
vignette: >
  %\VignetteIndexEntry{Scaling for Malaria Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


Load the required packages:

```{r}
library(ramp.xds)
library(ramp.library)
library(deSolve)
library(rootSolve)
library(ramp.work)
```

```{r, echo=F}
#devtools::load_all()
```

# Scaling

This is a dog.

## xde_scaling 

The function `xde_scaling()` defines the relationship between the EIR and the PR, and it outputs stable orbits for each value of aEIR in a mesh from $10^{-1}$ up to $10^{3}$ The code is in `mob_library/Work`
Do we need two versions?

The cohort trace functions in `ramp.xds` take the form of `F(a, bday=0, scale=1).` 

```{r}
F_sin = function(t){(1.01 + sin(2*pi*t/365))}
F_1 = function(t){0*t + 1/365}
```

```{r}
xds_setup_cohort(Xname = "SIS", F_season=F_1) -> sis
```

```{r}
xds_solve_cohort(sis) -> sis
```

```{r}
xde_scaling_eir(sis, 25) -> sis
```

```{r}
plot_eirpr(sis)
```

```{r}
require(viridis)
clrs = turbo(25)
```

```{r, fig.height=3.5, fig.width=5}
plot_eirpr(sis)

with(sis$output$eirpr,{
  points(aeir, pr, col = clrs)
  lines(scaling[[5]]$aeir, scaling[[5]]$pr, col = clrs[5])
  lines(scaling[[10]]$aeir, scaling[[10]]$pr, col = clrs[10])
  lines(scaling[[15]]$aeir, scaling[[15]]$pr, col = clrs[15])
  lines(scaling[[20]]$aeir, scaling[[20]]$pr, col = clrs[20])
})
```

## `xde_pr2eir()`

Since `xde_scaling` defines the relationship between the EIR and the PR, we can now run `xde_pr2eir()` to get the predicted value of the eir, for any given value of the pr. The code is in `mob_library/Work`


We can run this for 50 randomly chosen values of the *Pf*PR. 

```{r}
preir_i = xde_pr2eir(c(0.001, runif(25, 0, 1), 0.999), sis)
```

The function flags any values that are outside of the acceptable range. This may not seem important for the SIS model, but the range of other models can be bounded, so we don't want to return nonsense values.  

```{r}
preir_i$errors
```

We can plot the others:  

```{r}
plot_eirpr(sis)
with(sis$outputs$eirpr, points(aeir, pr, pch = 15))
with(preir_i, points(365*eir, pr, pch = 19, col = "red"))
```


## Scaling {.tabset}

### Seasonality

```{r}
sis0 <- xds_setup_cohort(Xname = "SIS", F_season = F_sin)
xde_scaling_eir(sis0, 25) -> sis0
```


```{r}
clrs = turbo(25)
with(sis$outputs$eirpr, plot(aeir, pr, type = "l", log = "x", xaxt= "n", xlab = "aEIR", ylab = "PR"))
axis(1, 10^(-1:3), c(0.1, 1, 10, 100, 1000))
lines(sis$outputs$eirpr$aeir, sis0$outputs$eirpr$pr, col = "tomato", lwd=2) 

with(sis0$outputs$eirpr, points(aeir, pr, col = clrs))
with(sis0$outputs$eirpr, lines(scaling[[5]]$aeir, scaling[[5]]$pr, col = clrs[5]))
with(sis0$outputs$eirpr, lines(scaling[[10]]$aeir, scaling[[10]]$pr, col = clrs[10]))
with(sis0$outputs$eirpr, lines(scaling[[15]]$aeir, scaling[[15]]$pr, col = clrs[15]))
with(sis0$outputs$eirpr, lines(scaling[[20]]$aeir, scaling[[20]]$pr, col = clrs[20]))
```

### Drug taking 

```{r}
sip = xds_setup_cohort(Xname = "SIP", F_season=F_sin)
sip$Xpar[[1]]$eta = 1/40
xde_scaling_eir(sip, 25) -> sip
```


```{r}
sip1 = setup_exposure_nb(sip, 1/50)
xde_scaling_eir(sip1, 25) -> sip1
```

```{r}
with(sis$outputs$eirpr, plot(aeir, pr, type = "l", log = "x", xaxt= "n", xlab = "aEIR", ylab = "PR"))
axis(1, 10^(-1:3), c(0.1, 1, 10, 100, 1000))
with(sip$outputs$eirpr, lines(aeir, pr, col = "darkorange"))
with(sip1$outputs$eirpr, lines(aeir, pr, col = "brown"))
``` 

### Environmental Heterogeneity 

```{r}
sis4 <- setup_exposure_nb(sis, 1/50)
xde_scaling_eir(sis4, 25) -> sis4
```

```{r}
with(sis$outputs$eirpr, plot(aeir, pr, type = "l", log = "x", xaxt= "n", xlab = "aEIR", ylab = "PR"))
axis(1, 10^(-1:3), c(0.1, 1, 10, 100, 1000))
#with(sis2$outputs$eir, lines(aeir, pr, col = "blue"))
#with(sis3$outputs$eir, lines(aeir, pr, col = "purple"))
with(sis4$outputs$eir, lines(aeir, pr, col = "darkblue"))
```

### Travel

```{r}
sis5 <- setup_travel_static(sis, delta = 1/5/365)
xde_scaling_eir(sis5, 25) -> sis5
```


```{r}
with(sis$outputs$eirpr, plot(aeir, pr, type = "l", log = "x", xaxt= "n", xlab = "aEIR", ylab = "PR"))
axis(1, 10^(-1:3), c(0.1, 1, 10, 100, 1000))
with(sis5$outputs$eir, lines(aeir, pr, col = "darkgreen")) 
```

