---
title: "Seasonality" 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document 
vignette: >
  %\VignetteIndexEntry{Seasonality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(ramp.xds)
#library(ramp.work)
devtools::load_all()
```

Here, we illustrate the use of functions that fit parameters for a seasonal pattern. 

## Pseudo Data

We generate a seasonal signal for the *Pf*EIR. 

```{r}
tt <- round(seq(0, 2500, by = 365/12)) 
Fpar <- makepar_F_sin(pw=3, floor=0.3, phase=172)
Fs0 <- make_function(Fpar) 
plot(tt/365, Fs0(tt), type = "l", xlab = "Time (Years)", ylab = "EIR")
```

```{r}
mod  <- xds_setup_cohort(1/365, F_season = Fs0)
mod  <- xds_solve_cohort(mod, times=tt)  
pr <- mod$outputs$orbits$XH[[1]]$true_pr 
```


```{r}
obs_pr <- pmax(0, rnorm(length(tt), pr, .08))
xds_plot_PR(mod)
lines(tt, obs_pr)
```

## Fit Phase and Amplitude

We create a new model that has no information about the seasonal pattern, but it has the right mean. 

```{r}
mod1 <- xds_setup_cohort(eir = 1/365) 
mod1 <- xds_solve_cohort(mod1, times=tt)
```

The first step is to fit the phase, using default parameter values for the amplitude. 

```{r}
fit_phase_sin_season(obs_pr, tt, mod1) -> mod1 
mod1$EIRpar$season_par$phase 
```

Next, we set the phase for a parameter set defining a seasonal pattern, and we fit the amplitude. We modify the function to have this fitted 

```{r}
mod1 <- fit_amplitude_sin_season(obs_pr, tt, mod1)
mod1$EIRpar$season_par
```

Note that if we pass this new parameter set to `fit_phase_sin_season,`  we get a slightly different phase. 

```{r}
fit_phase_sin_season(obs_pr, tt, mod1) -> mod1 
fit_amplitude_sin_season(obs_pr, tt, mod1) -> mod1
```

## Accuracy 

The algorithm fits the seasonal *Pf*PR pattern with reasonable accuracy. 

```{r}
mod1 <- xds_solve_cohort(mod1, times=tt)

xds_plot_PR(mod)
xds_plot_PR(mod1, clrs="darkred", add=TRUE)
lines(tt, obs_pr)
```


